Hierarchical Controller Architecture in Software Defined Networks

### 2013-12-14 paper/paper.tex
Related work section

\documentclass[10pt, twocolumn]{article}
\usepackage{url,graphicx,tabularx,array,geometry, caption, subcaption, float}

%-- Commands for header
\renewcommand{\title}[1]{\textbf{#1}\\}
\renewcommand{\line}{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}\hline\\\end{tabularx}\\[-0.10cm]}
\newcommand{\leftright}[2]{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}#1%
& #2\\\end{tabularx}\\[-0.10cm]}

@@

\begin{document}
\maketitle
\begin{abstract}
Network control plane, in Software Defined Networks (SDN), provides a global view of physical network to control applications (such as routing), which can be developed as a centralized application rather than a distributed one. Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple nodes called controllers, to achieve desired level of scalability, reliability and responsiveness.

Traditional architecture involves multiple controllers, each operating on a subset of switches and maintaining a global view of entire network. We propose a different where controllers are arranged in a hierarchy and only maintain a part of network view. This improves responsiveness and reduces inconsistency ........(numbers).

@@

\section{Introduction}
Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature. Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements. Some classes of control application are routing, traffic engineering (TE) and enforcing network policies. In traditional network, there is a tight coupling and one to one mapping between control elements and forwarding elements, with one pair on each switch. Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules. It is the control application which runs distributed protocol to synchronize the state of forwarding elements. The design of these distributed protocols differs from application to application. For example, link state routing makes switches share their neighbors while distance vector routing makes switches share their routing tables.

The emergence of Software Defined Network (SDN) has induced a separation between forwarding and control elements and has significantly improved the role of control plane. The control element may no longer be present on the switch but move to a separate device called controller, and uses predefined protocol such as OpenFlow [openflow-ref], to install forwarding rules. There may be a one-to-many mapping from controllers to switches. At one extreme, control plane may provide control application with global network view, which can then be implemented as a centralized application, rather than a distributed one. Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple controllers, to achieve desired level of scalability, reliability and responsiveness.

Conventional control plane architecture involves multiple controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application. Instructions (forwarding rules) from control application are sent to target switches directly or through other controllers. Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT). Consequently, this introduces trade-off between network state inconsistency and responsiveness. Results from [logically-centralized-ref] suggests that control state inconsistency significantly degrades performance of logically centralized control applications agnostic to the underlying state distribution.

The state inconsistency problem arises from the fact that each controller try to maintain the complete global view of network. Besides, we believe it is difficult (from scalability point of view) for a controller (a COTS machine) to handle the entire network view. Motivated by these observation, we propose a hierarchical controller architecture. We show ....................

@@

\section{Hierarchical Controller Architecture}
\label{sec:hierarchical}
We propose a control plane architecture in which controllers are arranged in the form of a tree over the physical network. Switches form the leaf nodes in the tree, while controllers form the internal nodes. An internal node acts as a switch for it's parent and acts as a controller for it's children. A node aggregate it's children network and converts it to a switch abstraction which is then exposed to the parent controller. This switch abstraction is explained in detail in section~\ref{subsec:aggr}. A controller's view does not span the whole network but only the network of it's children. A node translates instructions (forwarding rules) from it's parent controller into forwarding rules for it's underlying network. Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller. The translation process is explained in the following section.

@@

\subsection{An example}
\label{example}
Let's consider an example to understand the benefits of hierarchical architecture. Let there be a TE control application running on $d$ controllers, with each controller operating on $d$ switches. Network state consists of current utilizations of all links. Whenever a new flow arrives, application choses a path with minimum maximum utilization and assigns it to the flow. Each time a path of length $l$ is assigned to the flow, there could be a maximum of $l*d$ link utilization updates and $l$ forwarding rule messages to be sent between controllers. Now, consider an architecture where controllers are arranged in hierarchical manner. There are $d+1$ controllers. The first $d$ controllers (called type 0) maintain $d$ switches each and acts as a switch to the last controller (called type 1), which in turn maintains the other $d$ controllers. Type 0 controllers keep track of link utilizations only for the links that connect it's children switches. Type 1 controller keep track of links that runs from switches belonging to one controller domain to another. Same TE control application runs on each controller. When a new flow arrives, a switch sends $packet_in$ message to it's type 0 controller. Type 0 controller checks if the path of the flow belongs entirely to it's underlying network or not. If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller. However, if the destination is outside it's domain, type 0 controller sends a $packet_in$ message to it's parent controller. Parent controller uses TE control application to assign a paths of type 0 controllers. Forwarding rules are sent to type 0 controllers, which in turn sets up forwarding rules in their network. If the path length is $l$, there will be a maximum of $l+1$ forwarding rule messages between controllers and no link utilization updates. This significant reduction in communication overhead comes at the cost of sub-optimality in the application performance. We show experimentally that the proposed architecture significantly reduces communication overhead while only modestly degrading application performance in the next section.

@@

\section{Related Work}
\label{sec:related}

\section{Conclusion}
\label{sec:conc}

>>>>

\documentclass[10pt, twocolumn]{article}
\usepackage{url,graphicx,tabularx,array,geometry, caption, subcaption, float}
%-- Commands for header
\renewcommand{\title}[1]{\textbf{#1}\\}
\renewcommand{\line}{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}\hline\\\end{tabularx}\\[-0.10cm]}
\newcommand{\leftright}[2]{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}#1%
& #2\\\end{tabularx}\\[-0.10cm]}

@@

\begin{document}
\maketitle
\begin{abstract}
Network control plane, in Software Defined Networks (SDN), provides a global view of the physical network to control applications (such as routing), which can be developed as a centralized application rather than a distributed one. Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple nodes called controllers, to achieve desired level of scalability, reliability and responsiveness.

Traditional architecture involves multiple controllers, each operating on a subset of switches and maintaining a global view of the entire network. We propose a different architecture where controllers are arranged in a hierarchy and only view a part of the network. This improves responsiveness and reduces inconsistency ........(numbers).

@@

\section{Introduction}
Forwarding or data plane consists of a collection of forwarding elements (switches), connected to form a network, which is inherently distributed in nature. Control plane consists of a collection of control elements, which install forwarding rules on forwarding elements as dictated by control logic or application, running on control elements. Some classes of control applications are routing, traffic engineering (TE), enforcing network policies, etc. In traditional networks, there is a tight coupling and a one to one mapping between control elements and forwarding elements, with one pair on each switch. Besides, the role of control elements had been trivial, which is to merely install instructions given by control application on corresponding forwarding element as forwarding rules. It is the control application which runs distributed protocols to synchronize the state of forwarding elements. The design of these distributed protocols differs from application to application. For example, link state routing makes switches share their neighbors while distance vector routing makes switches share their routing tables.

The emergence of Software Defined Networks (SDN) has induced a separation between forwarding and control elements and has significantly improved the role of the control plane. The control elements may no longer be present on the switches but move to a separate device called a controller, and use predefined protocol such as OpenFlow [openflow-ref], to install forwarding rules. There may be a one-to-many mapping from controllers to switches. At one extreme, control plane may provide a control application with the global network view, which can then be implemented as a centralized application, rather than a distributed one. Regardless of this abstraction, the control plane itself must inevitably be a distributed system, with multiple controllers, to achieve desired level of scalability, reliability and responsiveness.

Conventional control plane architectures involve multiple controllers, where each controller operates on a subset of switches and maintains a global view of the entire network, which is exposed to the control application. Instructions (forwarding rules) from control application are sent to target switches directly or through other controllers. Underlying network state is shared with other controllers through distributed state dissemination frameworks such as distributed transactional databases or Distributed Hash Tables (DHT). Consequently, this introduces a trade-off between network state inconsistency and responsiveness. Results from [logically-centralized-ref] suggests that control state inconsistency significantly degrades performance of logically centralized control applications agnostic to the underlying state distribution.

The state inconsistency problem arises from the fact that each controller tries to maintain the complete global view of the network. Besides, we believe that it is difficult (from scalability point of view) for a controller (a COTS machine) to handle the entire network view. Motivated by these observations, we propose a hierarchical controller architecture. We show ....................

@@

\section{Hierarchical Controller Architecture}
\label{sec:hierarchical}
We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network. Switches form the leaf nodes in the tree, while controllers form the internal nodes [TODO:tree-figure]. An internal node acts as a switch for its parent and as a controller for its children. A node aggregates its underlying network and converts it to a switch abstraction which is then exposed to the parent controller. This switch abstraction is explained in detail in section~\ref{subsec:aggr}. A controller's view does not span the whole network but only the network of its children. A node translates instructions (forwarding rules) from its parent controller to the forwarding rules for its underlying network. Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller. The translation process is explained in the following section.

@@

\subsection{An example}
\label{example}
Let's consider an example to understand the benefits of hierarchical architecture. Let there be a TE control application running on $d$ controllers, with each controller operating on $d$ switches. Network state consists of current utilizations of all links. Whenever a new flow arrives, application choses a path with minimum maximum utilization and assigns it to the flow. Each time a path of length $l$ is assigned to the flow, there could be a maximum of $l*d$ link utilization updates and $l$ forwarding rule messages to be sent between controllers. Now, consider an architecture where controllers are arranged in hierarchical manner. There are $d+1$ controllers. The first $d$ controllers (called type 0) maintain $d$ switches each and acts as a switch to the last controller (called type 1), which in turn maintains the other $d$ controllers. Type 0 controllers keep track of link utilizations only for the links that connect it's children switches. Type 1 controller keep track of links that runs from switches belonging to one controller domain to another. Same TE control application runs on each controller. When a new flow arrives, a switch sends $packet_{in}$ message to its type 0 controller. Type 0 controller checks if the path of the flow belongs entirely to its underlying network or not. If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller. However, if the destination is outside its domain, type 0 controller sends a $packet_{in}$ message to its parent controller. Parent controller uses TE control application to assign a paths of type 0 controllers. Forwarding rules are sent to type 0 controllers, which in turn sets up forwarding rules in their network. If the path length is $l$, there will be a maximum of $l+1$ forwarding rule messages between controllers and no link utilization updates. This significant reduction in communication overhead comes at the cost of sub-optimality in the application performance. We show experimentally that the proposed architecture significantly reduces communication overhead while only modestly degrading application performance in the next section.

@@

\section{Related Work}
\label{sec:related}
Owing to the performance, reliability and scalability issues faced by single controller architectures, distributed controller architectures have been studied and implemented in various flavors. Most of them aim for a logically centralized global view of the network, while Some of them are discussed in this discussion.

Hyperflow[Hyperflow-ref] has a logically centralized but physically distributed controller architecture. It localizes decision making in individual controllers by passively synchronizing the global network view. The control application is agnostic to the underlying architecture and takes a global decision, though the effect is local. The events are propagated using a publish/subscribe messaging paradigm.

Onix[Onix ref] presents a platform on top of which a network control plane can be implemented as a distributed system. Control planes written within Onix operate on a global view of the network and use basic state distribution primitives provided by the platform. Thus Onix provides a general distributed state management API for control plane implementations, but leaves the decision making and trade-offs among consistency, durability and scalability to the control applications. This makes the design, deployment, modification and verification of control applications difficult and cost prohibitive.

Kandoo[Kandoo-ref] takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state. Bottom layer runs a local application for its underlying network, while the top layer subscribes to changes from the bottom layer. Top layer view is subscription dependent and the architecture cannot be used in a generic fasion.

On a similar but different note, B4[B4-ref] uses a two level hierarchy, where the site controllers view their local network and aggregate flows to the higher level centralized controller which performs TE. Though the paper talks about a very particular implementation, it demonstrates in practice the benefits of the generic principle of our design.
\section{Conclusion}
\label{sec:conc}
